The effects of priming in human computation

### 2014-05-14 docs/publications/2014.05.9_pnas/pnas2.tex
redo discussion of statistics

We can then calculate the relative specificity of two treatments to be the
average relative specificity of workers drawn uniformly from the treatments:

\begin{align}
	S(U,V) = 
		\frac{1}{|U \times V|}
		\sum_{(u,v) \; \in \; U \times V} \;
		s(u,v),
		\label{eq:specificity}
\end{align}

where $U$ and $V$ are treatments, taken as sets of workers. \textbf{Figure 9} 
exhibits the results.

@@

\textbf{Building the ontology of terms.}  The ontology was built manually
by the experimenters, starting from the corpus of labels. Labels occuring only 
once were discarded.  We created the root classes \texttt{activity}, 
\texttt{adjective}, \texttt{culture}, 
\texttt{food}, and \texttt{thing}. \texttt{culture} and \texttt{food} were 
given top-level status because they were prominently represented in the corpus.
Whenever one label was judged be regarded as a more specific case of another, a
directed edge was drawn from the more general to the more specific. This was a 
judgment call was made by the experimenters.  However, the labels from all 
treatments were aggregated before assembling the ontology.  This was done to 
ensure that the construction of the ontology would not be affected by 
information about the treatments from which a label originated. 
Labels which were judged to be roughly synonymous were treated as identical,
and labels judged to be misspellings were corrected.

\textbf{Statistical significance for relative specificity.}
We used \textbf{Eq.~\ref{eq:specificity}} to compute the relative specificity
of two treatments.  To analyze relative specificity statistically,
we first provide an equivalent rephrasing of \textbf{Eq.~\ref{eq:specificity}}.
Let the specificity of a worker $u$ relative to a 
\textit{basis treatment} $V$, denoted $S_V(u)$, be the expected relative 
specificity between $u$ and a worker uniformly sampled from $V$:

\begin{align}
	S_V(u) =  \text{E}\{s(u,v)\}, \quad v \in_R V.
\end{align}
Then, the relative specificity of two treatmens, $S(U,V)$, can be defined as:
\begin{align}
	S(U,V) = \text{E}\{S_V(u)\} \quad u \in_R U.
\end{align}

We can estimate this using the sample mean calculated from our treatments:
\begin{align}
	\hat{S}_V(u) &= \frac{1}{|V|} \sum_{v \in V} s(u,v), \\
	\hat{S}(U,V) &= \frac{1}{|U|} \sum_{u \in V} S_V(u), \\
	\text{Var}\{\hat{S}_V(u)\} &= \frac{\text{Var}\{s(u,v)\}}{|V|},\\
	\text{Var}\{\hat{S}(U,V)\} &= \sum_{u\in U}\text{Var}\{\hat{S}_V(u)\}
\end{align}

And rewrite the relative specificity of $U$ and $V$ as the expected value of
$S_V(u)$ for workers in $U$:
\begin{align}
	S(U,V) &= E\{S_V(u)\}, \\
	\text{with} \quad\quad \text{Var}\{S(U,V)\} 
		&= \frac{\text{Var}\{S_V(u)\}}{|U|}.
\end{align}
To decide if $U$ is significantly more (less) specific than $V$, we do
a null comparison of $V$ to itself: $S(V,V)$.  This value is in principle zero,
but the comparison yields $\text{Var}\{S(V,V)\}$ which we use to state and 
test the null hypothesis that $U$ and $V$ are equally specific:
\begin{align}
	\mathbf{H_0}: \quad S(U,V)-S(V,V)=0
\end{align}
Note that since $S(V,V)=0$, $\mathbf{H_0}$ essentially states that $S(U,V)=0$,
however, it is more stringent since we consider variance due to sampling both
$U$ and $V$:
\begin{align}
\text{Var}\{S(U,V) - S(V,V)\} &= \text{Var}\{S(U,V)\} + \text{Var}\{S(V,V)\}\\
						&> \text{Var}\{S(U,V)\}
\end{align}
Therefore, we find $U$ to be significantly more (less) specific than $V$ when:
\begin{align}
z = \frac{S(U,V)}{ \sqrt{ \text{Var}\{S(U,V)\} + \text{Var}\{S(V,V)\} }} 
	> 1.96.
\end{align}

Due to 
computational intensity, we randomly subsampled 63 workers from each treatment.
To compute the relative food- or culture-specificity, we alter the calculation
in \textbf{Eq.~\ref{eq:worker-specificity}} by only including labels 
respectively of food- or culture-orientation, while excluding all labels that
aro both food- and culture-oriented.

>>>>

We can then define the relative specificity of two treatments, 
$S(\mathcal{U},\mathcal{V})$ to be the average
relative specificity of workers drawn uniformly from the treatments.  
$\hat{S}({\mathcal{U}, \mathcal{V}})$ estimates this using the sample mean:

\begin{align}
	\hat{S}(\mathcal{U},\mathcal{V}) = 
	\frac{1}{|\mathcal{U} \times \mathcal{V}|}
	\sum_{(u,v) \; \in \; \mathcal{U} \times \mathcal{V}} \;
		s(u,v),
		\label{eq:specificity}
\end{align}

where $\mathcal{U}$ and $\mathcal{V}$ are treatments, taken as sets of workers.
See \textbf{Materials and Methods} for statistical analysis of the relative 
specificity of two treatments. 
\textbf{Figure 9} 
exhibits the results.

@@

\textbf{Building the ontology of terms.}  The ontology was built manually
by the experimenters, starting from the corpus of labels. Labels occuring only 
once were discarded.  We created the root classes \texttt{activity}, 
\texttt{adjective}, \texttt{culture}, 
\texttt{food}, and \texttt{thing}. \texttt{culture} and \texttt{food} were 
given top-level status because they were prominently represented in the corpus.
Whenever one label was judged to be a more specific case of another, a
directed edge was drawn from the more general to the more specific. This
judgment call was made by the experimenters.  However, the labels from all 
treatments were aggregated before assembling the ontology,  to 
ensure that the ontology construction would not be affected by 
knowledge of the treatments from which labels originated. 
Labels which were judged to be roughly synonymous were treated as identical,
and labels judged to be misspellings were corrected.

\textbf{Statistical significance for relative specificity.}
We used \textbf{Eq.~\ref{eq:specificity}} to compute the relative specificity
of two treatments.  To analyze relative specificity statistically,
we first develop \textbf{Eq.~\ref{eq:specificity}} from an intermediate 
definition.  Let the specificity of a worker $u$ relative to a 
\textit{basis treatment} $\mathcal{V}$, denoted $S_V(u)$, to be the expected 
relative specificity between $u$ and a worker uniformly sampled from 
$\mathcal{V}$:

\begin{align}
	S_V(u) =  \text{E}_V\{s(u,V)\}, \quad V \in_R \mathcal{V},
		\label{eq:basis-specificity}
\end{align}
where $V$ is a random, uniformly sampled worker from $\mathcal{V}$.  We 
subscript the expectation operator with $V$ to indicate 
that the expectation is taken over the distribution of $V$.  Then, the relative specificity of two treatmens, $S(\mathcal{U},{V})$, can be 
written as:
\begin{align}
	S(\mathcal{U},\mathcal{V}) = 
		\text{E}_U\{S_V(U)\}, \quad U \in_R \mathcal{U},
		\label{eq:treatment-specificity}
\end{align}
where $U$ is a uniformly sampled worker from $\mathcal{U}$.  This yields 
estimators for $S_V(u)$ and $S(\mathcal{U}, \mathcal{V})$, and their variances
as shown below (note that 
\textbf{Eq.~\ref{eq:treatment-specificity-estimator}} is
equivalent to \textbf{Eq.~\ref{eq:specificity}}):

\begin{align}
	\hat{S}_V(u) &= \frac{1}{|\mathcal{V}|} \sum_{v \in \mathcal{V}} s(u,v), 
		\label{eq:basis-specificity-estimator} \\
		\hat{S}(\mathcal{U},\mathcal{V}) 
			&= \frac{1}{|\mathcal{U}|} \sum_{u \in \mathcal{V}} S_V(u),
		\label{eq:treatment-specificity-estimator} \\
		\text{Var}_V\{\hat{S}_V(u)\} 
			&= \frac{1}{|\mathcal{V}|}\text{Var}_V\{s(u,V)\},
		\label{eq:basis-specificity-variance} \\
		\text{Var}_{U,V}\{\hat{S}(\mathcal{U},\mathcal{V})\} 
			&= \frac{1}{|\mathcal{V}|\cdot|\mathcal{U}|^2} 
			\sum_{u\in U}\text{Var}_V\{{s}(u,V)\}.
		\label{eq:treatment-specificity-variance}
\end{align}
$\text{Var}_V{s(u,V)}$ can be directly calculated as a sample variance, so
\textbf{Eq.~\ref{eq:basis-specificity-estimator}} through \textbf{\ref{eq:treatment-specificity-variance}} provide enough information to test the hypothesese:
\begin{align}
	\begin{matrix}
		\mathbf{H_0}: S(\mathcal{U}, \mathcal{V}) = 0 \\[0.5em]
		\mathbf{H_1}: S(\mathcal{U}, \mathcal{V}) \neq 0
	\end{matrix}
\end{align}
